On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-02 main.tex
Thom

\usepackage{mathrsfs}
% new commands and shortcuts
\def\bs{\textit{\textbf{s}}}
\def\Lb{\textit{\textbf{L}}}
\def\Xb{\textit{\textbf{X}}}
\def\xb{\textit{\textbf{x}}}
\def\fb{\textit{\textbf{f}}}
\def\ab{\textit{\textbf{a}}}
\def\ajb{\overline{a_j}}
\def\bjb{\overline{b_{j,l}}}
\def\pjb{\overline{P_{j}}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\htt}{ht}
\def\C{\mathbb{C}}
\def\vt{\vartheta}
\def\Q{\mathbb{Q}}
\def\R{\mathbb{R}}
\def\rs{\mathscr{R}}
\def\P{\mathscr{P}}
\def\d{\delta}
\def\sing{ \textrm{sing}}
\def\codim{ \textrm{codim}}
\def\jac{ \textrm{jac}}
\def\exp{\textrm{exp}}
\def\grad{\textrm{grad}}
\def\rank{\textrm{rank}}
\def\reg{\textrm{reg}}
\def\jt{\widetilde{J}}
\def\At{\widetilde{A}}
\def\Yt{\widetilde{Y}}
\def\dt{\widetilde{d}}
\def\Dt{\widetilde{D}}
\def\rk{\textrm{rank }}
\def\pa{\partial}
\def\D{\Delta}
\def\Z{\frak{Z}}
\def\I{\frak{I}}
\def\Is{\frak{I}^{\star}}
\def\A{\frak{A}}
\def\fp{\frak{P}}
\def\sp{\mathscr{P}}
\def\la{\langle}
\def\ra{\rangle}
% matrices 
\def\bbm{\begin{bmatrix}}
\def\ebm{\end{bmatrix}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

  % Furthermore, as bit complexity provides a far more realistic model
  % of computation compared to arithmetic complexity, which only
  % counts each operation at unit cost whereas bit complexity provide
  % more information for both time and space resources that are needed
  % in practice, it is important that this missing gap is filled.
  % As bit complexity provides a far more realistic model of
  % computation compared to arithmetic complexity, which only counts
  % each operation at unit cost, our contribution is necessary and
  % important. Indeed, bit estimates provide more information for both
  % time and space resources that are needed in practice.
\end{abstract}

@@

Write more about transversality? Provide an example?
\subsection{Extending Thom's weak transversality}
We prove a quantitative extension of Thom's weak transversality \cite{TWT} specialized to the particular case of transversality to a point, which can be rephrased entirely in terms of critical / regular values.
Denote by $\Phi :\C^n \times \C^{\dt} ~ \rightarrow \C^{m}$
a polynomial mapping, where $n,\dt,$ and $m$ are positive integers (assume $m \leq n$?). Assume the total degree of $\Phi$ is bounded by an integer $d$. For $\vt$ in $\C^{\dt}$, let $\Phi_{\vt} : \C^n \rightarrow \C^{m}$ be the induced mapping $x\mapsto \Phi(x,\vt)$.
\begin{theorem} (Weak transversality)
Suppose that $0$ is a regular value of $\Phi$. Then there exists a hypersurface $\Delta \subset \C^{\dt}$ of degree at most $d^n$ for which, if $\vt \in \C^{\dt}-\Delta$ then $0$ is a regular value of $\Phi_{\vt}$. 

@@

It will sometimes be useful for us to let the matrix $A$ represent new indeterminates: let $\A$ be an $n \times n$ matrix of new indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in \mathbb{Q}(\A_{i,j})[\textbf{X}]$ as $f(\A\textbf{X})$ and $V^\A=V(f^\A).$ It follows from Proposition 4.2 that for a generic $A \in \C^{n^2},$ $0$ is a regular value of $\Phi_{A}$.  The proof of \cite[Theorem B.3]{NO}, which shows the existence of $\Delta \subset \C^{n^2},$ applies \textit{Sard's lemma}, which shows that the critical values of a polynomial mapping are contained in a hypersurface. It follows from Sard's lemma that for the matrix of indeterminants $\A,$ $0$ is a regular value of $\Phi_{\A}.$ Otherwise it could not be that $0$ is a regular value of $\Phi_A,$ for a generic $A\in \C^{n^2}.$ Therefore,

@@

has full rank for all $\xb \in V(f^{\A})$, and thus $f^{\A}$satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$. We will rely on this fact in Section 5. 

@@

\subsection{Proof of Proposition 5.1}
The proof of \cite[Theorem B.3]{TWT} shows the existence of $\Delta.$ We extend the theorem by showing that the degree of $\Delta$ is upper bounded by $d^n.$ 
\subsubsection{More notation and setup}
Put $X = \Phi^{-1}(0)$ and consider the projection $\pi:(x, \vt) \in \C^n \times \C^{\dt} \mapsto \vt \in\C^{\dt}$. The proof of \cite[Theorem B.3]{NO} also shows that $X$ is $(n+ \dt -m)$-equidimensional, and that if $\vt \in \C^{\dt}$
is such that $0$ is a critical value of $\Phi_{\vt}$, then $\vt$ is a critical value of the restriction of $\pi$ to $X.$ We show that the critical points of $\pi|_X$ are contained in an algebraic set
$\Delta^{'}$. We then show that 
\deg \Delta \leq \deg \Delta^{'} \leq d^n.
\subsubsection{Explicitly characterizing $\Delta^{'}$.} 
We will use \cite[Lemma B.4]{NO}:
\begin{lemma}
Let $A=\bbm A_1 \\ A_2 \ebm$ be a matrix. Then, 
\textup{rank}(A) = 
\textup{rank}(A_1) + \textup{rank}(A_2|\ker(A_1)),
where $A_2|\ker(A_1)$ is the restriction of the linear map defined by $A_2$ to the kernel of $A_1$. 
\end{lemma}
Let $M$ denote the matrix
M(\xb,\vt)&:= \jac_{(\xb,\vt)}(\Phi;X)\\
&= 
\jac_{(\xb,\vt)}(\pi)\\
\jac_{(\xb,\vt)}(\Phi) 
\ebm 
=
\textbf{0}_{\dt \times n}\hspace{5mm}\textbf{I}_{\dt} \\
\jac_{(\xb,\vt)}(\Phi)
\ebm.
$M(\xb,\vt)$ has full rank $m+\dt$ if and only if $(\xb,\vt)$ is a regular point on $\pi|_X.$ 
The dimension of the source $\pi|_X$ is $\dim X = n + \dt - m.$ The dimension of the target space is $\dim \C^{\dt} = \dt.$ Now, if $(\xb,\vt)$ is a regular point of $\pi|_{X}$ then $\dim \textrm{Im}(d\pi|_X(T_{\xb}X)) = \dt$. When $\xb$ is regular on $X$, by Proposition 2.3, $\jac_{(\xb,\vt)} (\Phi) = T_{(\xb,\vt)}X$ and has full $\rk n + \dt - \dim X = n + \dt - (n + \dt - m) = m$. By applying Lemma 4.3 we have
\rank ~M(\xb,\vartheta) &= 
\rank ~
\jac_{(\xb,\vartheta)}(\pi|_X)\\
\jac_{(\xb,\vartheta)}(\Phi) 
\ebm \\
&= \rank~ \jac_{(\xb,\vartheta)}(\pi|_X) + \rank ~\jac_{(\xb,\vartheta)}(\Phi)|_{T_{(x,\vartheta)}X} \\
&= m + \dim \textrm{Im}(d\pi|_{\Xb}(T_xX)) \\
&= m + \dt 
Otherwise, if $(\xb,\vt)$ is not a regular point of $\pi|_X,$ then \[\dim \textrm{Im}(d\pi|_X(T_{\xb}X)) < \dt\]  and therefore $\rank ~M(\xb,\dt) < m+\dt.$ 
Therefore if $M(x,\vt)$ does not have full rank $\dt+m$ then $(x,\vt)$ is a critical point of $\pi|_X$. Since the minors of order $\dt+m$ vanish when $(x,\vt)$ is a critical point, the algebraic set defined by the vanishing of the minors of $M$ of order $\dt + m$ describe the critical points:
\Delta^{'} = V(\textrm{minors of }M\textrm{ of order }\dt + m).
Furthermore, we can characterize $\Delta^{'}$ with fewer equations by discarding those that evaluate to zero. Indeed, the first $\dt$ rows of $M(x,\vt)$ is the Jacobian matrix of the projection mapping $\pi|_X,$ which consists of a sub-matrix of zeros followed by an identity sub-matrix. Hence, we can discard several minors that evaluate to zero. Let $J$ denote the sub-matrix of the Jacobian of $\Phi$ consisting of the first $n$ columns:
 J := \jac_{(\xb,\vt)}(\Phi)[-;1,n].
$J(\xb,\vt)$ has full rank $\min\{m,n\}$ if and only if $M(\xb,\vt)$ has full rank $\dt+m.$ 
Notice 
M(\xb,\vt)&= 
\jac_{(\xb,\vt)}(\pi)\\
\jac_{(\xb,\vt)}(\Phi) 
\ebm \\
&=
\textbf{0}_{\dt \times n}\hspace{5mm}\textbf{I}_{\dt} \\
\jac_{(\xb,\vt)}(\Phi)
\ebm\\
&=
\textbf{0}_{\dt \times n} &\textbf{I}_{\dt} \\
J(\xb,\vt)     &\jac_{(\xb,\vt)}(\Phi)[-;n+1,\dt]
\ebm.
The first $n$ columns of $M(\xb,\vt)$ must all be different from the remaining $\dt$ columns. Therefore $J(\xb,\vt)$ has full rank $\min\{n,m\}$ if and only if $M(\xb,\vt)$ has full rank $\dt+m.$
Therefore $\Delta^{'} = V(\textrm{minors of }J\textrm{ of order }\min\{m,n\}).$
\subsubsection{Bounding the degree of $\Delta^{'}$.}
Let $\textit{\textbf{L}}=[L_1,\hdots,L_m]$ be Lagrange multipliers, and consider the Lagrange system 
\textit{\textbf{L}}\cdot J(x,A)=[\mathscr{L}_1,\hdots,\mathscr{L}_{n}]=[0,\hdots,0].
Denote by $\frak{Z}$ the algebraic set defined by the vanishing of $\mathscr{L}_1,\hdots,\mathscr{L}_{n},$ and consider the projection 
\pi_{n+\dt} :~ \C^{n+\dt+m} &\rightarrow \C^{n+\dt}\\
(\textit{\textbf{x}},\vt,\textit{\textbf{L}})~ &\mapsto (\textit{\textbf{x}},\vt).
Denote by $\frak{Z}^{'}$ the algebraic set
\frak{Z}^{'} := \overline{\frak{Z} - \{(\textit{\textbf{x}},\vt,0,\hdots,0) \in \C^{n+\dt+m}~|~(\textit{\textbf{x}},\vt,0\hdots,0) \in \frak{Z}\}}.
The projection $\pi_{n+\dt}$ on $\frak{Z}^{'}$ is equal to $\Delta^{'}$.
We will use the following lemma to prove Proposition 4.6.
\begin{lemma} 
For an irreducible component $Z$ of $\Z^{'},$ there exists a Zariski open subset $Z^o \subset Z$ with \[\pi_{n+\dt}~(Z^{o}) \subset \D^{'}~(\textrm{why?})\]
\overline{\pi_{n+\dt}~(\overline{Z^o})} = \overline{\pi_{n+\dt}~(Z^o)}.
\end{lemma} 
There exists a Zariski open subset $Z^o \subset Z$ with 
\pi_{n+\dt}~(Z^{o}) \subset \D^{'}~(\textrm{why?}).
Since $Z^o \subset \overline{Z^o}$, we have $\pi_{n+\dt}~(Z^o) \subset \pi_{n+\dt}~(\overline{Z^o})$ and thus 
\overline{\pi_{n+\dt}~(Z^o)} \subset \overline{\pi_{n+\dt}~(\overline{Z^o})}.
Since closed sets are by definition varieties in the Zariski topology, \[\overline{\pi_{n+\dt}~(\overline{Z^o})} = V\left( I(\overline{Z^o}) \cap \mathbb{C}[\textit{\textbf{X}},\vt]\right),
so let 
\overline{\pi_{n+\dt}~(\overline{Z^o})} = V(\mathscr{I}),
for some ideal $\mathscr{I} \subset \C[\Xb,\vt].$ Take $f$ in $\mathscr{I}$ so that $f$ vanishes on 
\overline{\pi_{n+\dt}~(Z^o)}\subset \overline{\pi_{n+\dt}~(\overline{Z^o})}
Then $f$ must also vanish on the subset 
\pi_{n+\dt}~(Z^o)\subset \overline{\pi_{n+\dt}~(Z^o)}\subset \overline{\pi_{n+\dt}~(\overline{Z^o})}
and $Z^o,$ and hence also on $\overline{Z^o}.$ Therefore, 
f \in I(\overline{Z^o}) \cap \mathbb{}[\textit{\textbf{X}},\vt]
and  $\mathscr{I} \subset I(\overline{Z^o}) \cap \mathbb{Q}[\textit{\textbf{X}},\vt]$ (why?), and it then follows that 
\overline{\pi_{n+\dt}~(\overline{Z^o})}&=V\left(I(\overline{Z^o})\cap \mathbb{C}[\textit{\textbf{X}},\vt]\right) \\&\subset V(\mathscr{I})\\&\subset \overline{\pi_{n + \dt}~(Z^o)}\textrm{ (why?)}.
Now we prove Proposition 4.7.
For all $(\textit{\textbf{x}},\vt) \in \D^{'}, $ there exists $\textit{\textbf{L}}^{'} \not = (0,\hdots,0)$ with $(\textit{\textbf{x}},\vt,\textit{\textbf{L}}^{'}) \in \Z.$ Thus, $(\textit{\textbf{x}},\vt,\textit{\textbf{L}}^{'})$ is in
     \Z - \left\{(x,\vt,0,\hdots,0) \in \C^{n+\dt+m}~|~(x,\vt,0\hdots,0) \in \Z\right\}, 
which implies that $(x,\vt,\textit{\textbf{L}}^{'}) \in \Z^{'}$, which then in turn implies that
    (x,\vt) \in \Pi_{n + \dt}~(\Z^{'}).
Therefore $\D^{'} \subset \pi_{n+\dt}~(\Z^{'}).$
\newline \indent 
Now let $Z$ be an irreducible component of $\Z^{'}.$ By Lemma 4.7 there exists a Zariski open subset $Z^o \subset Z$ with $\pi_{n+\dt}~(Z^{o}) \subset \D^{'}$, and
 \Pi_{n+\dt}~(Z) = 
  \Pi_{n+\dt}~(\overline{Z^o})
  &\subset 
  \overline{\pi_{n+\dt}~(\overline{Z^o})}\\
  &=
  \overline{\Pi_{n+\dt}~(Z^o)}, &\textrm{by Lemma 4.7,}\\
  &\subset \overline{\D^{'}}=\D^{'}.
   Therefore $\Pi_{n+\dt}~(\Z^{'}) \subset \D^{'}.$
The degree of $\D^{'}$ is at most $d^n.$
It follows from B\'ezout's Theorem that 
\deg \frak{Z} \leq (d-1+1)^n=d^n.
From \cite[Lemma 2]{H} we have 
\deg(\pi_{n+\dt}(\Z^{'})) \leq \deg(\Z^{'}).
Therefore 
\deg(\D^{'}) &= \deg(\pi_{n+\dt}~(\Z^{'})),&\textrm{by Proposition 4.6,}\\ 
            &\leq \deg(\Z^{'}),\\ 
            &\leq \deg(\Z), &\textrm{by ?,}\\ 
            &\leq d^n.
\section{Proof of $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$}

@@

If the Jacobian of $\fb$ has full rank at all $x \in V(f) \cap \mathbb{R}^n,$ then it follows by the Jacobian criterion [4, Theorem 16.19] that $f$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$. 

@@

There exists a Zariski closed set $\Delta \subset \C^{n^2}$ of degree at most $D^n$ with the property that, if $A \in \C^{n^2} - \Delta$ then for each $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$.

@@

has full rank for all $\xb \in V(f^A)$. Then, by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$.

@@

has full rank for all $\xb \in V(f^A)$ ($\cap \R^n$?). Therefore, by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$, for each $i \in \{1,\hdots,n\}.$

@@

\section{Proof of $\textbf{H}_i^{'}(4),\textbf{H}_i^{'}(5)$ and $\textbf{H}_i^{'}(6)$}

@@

\section{Proof of $\textbf{H}_i^{'}(7)$}

>>>>

% test 
\usepackage{mathrsfs}
\usepackage{bm}
% new commands and shortcuts
\def\td{{\bf todo}}
\def\bs{\textit{\textbf{s}}}
\def\Lb{\textit{\textbf{L}}}
\def\Xb{\textit{\textbf{X}}}
\def\Lambdab{\bm{\Lambda}}
\def\Thetab{\bm{\Theta}}
\def\thetab{\bm{\vartheta}}
\def\xb{\textit{\textbf{x}}}
\def\fb{\textit{\textbf{f}}}
\def\ab{\textit{\textbf{a}}}
\def\ajb{\overline{a_j}}
\def\bjb{\overline{b_{j,l}}}
\def\pjb{\overline{P_{j}}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\htt}{ht}
\def\C{\mathbb{C}}
\def\vt{\vartheta}
\def\Q{\mathbb{Q}}
\def\R{\mathbb{R}}
\def\rs{\mathscr{R}}
\def\P{\mathscr{P}}
\def\d{\delta}
\def\sing{ \textrm{sing}}
\def\codim{ \textrm{codim}}
\def\jac{ \textrm{jac}}
\def\exp{\textrm{exp}}
\def\grad{\textrm{grad}}
\def\rank{\textrm{rank}}
\def\reg{\textrm{reg}}
\def\jt{\widetilde{J}}
\def\At{\widetilde{A}}
\def\Yt{\widetilde{Y}}
\def\dt{s}
%% \def\dt{\widetilde{d}}
\def\Dt{\widetilde{D}}
\def\rk{\textrm{rank }}
\def\pa{\partial}
\def\D{\Delta}
\def\Z{\frak{Z}}
\def\I{\frak{I}}
\def\Is{\frak{I}^{\star}}
\def\A{\frak{A}}
\def\fp{\frak{P}}
\def\sp{\mathscr{P}}
\def\la{\langle}
\def\ra{\rangle}
% matrices 
\def\bbm{\begin{bmatrix}}
\def\ebm{\end{bmatrix}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

\end{abstract}

@@

Sard's lemma states that the set of critical values of a smooth
function $\R^n \to \R^m$ has measure zero. One can give ``algebraic''
versions of it, for semi-algebraic mappings $\R^n \to \R^m$ as
in~\cite[Chapter~9]{bochnak1998real} or polynomial mappings $\C^n \to
\C^m$ as in~\cite[Chapter~3]{Mumford76}, for which the sets of critical
values are contained in strict semi-algebraic, resp.\ algebraic sets
in the codomain. Thom's weak transversality lemma, as given for
instance in~\cite{demazure2000bifurcations}, generalizes Sard's
lemma. In this section, we consider a particular case of this result
(transversality to a point), and establish a quantitative version of
it. Our presentation is inspired by that in~\cite{TWT}.

%% ; it states that if a smooth function
%% $F(\Xb,\Lambda)$, depending on parameters $\Lamdba$ in $\C^s$ is
%% ``transverse'' to a certain submanifold in its domain space, then for
%% most values $\lambda \in \C^s$, $F(\Xb,\lambda)$ remains

Transversality to a point can be rephrased entirely in terms of
critical and regular values. Recall that if $\bm \Psi$ is a mapping
from a smooth algebraic set $Y$ to $\C^t$, with $t\le \dim(Y)$, a {\em
  critical point} of $\bm \Psi$ is a point $\bm y \in Y$ such that the
image of the tangent space $T_{\bm y} Y$ by the differential $d_{\bm
  Y} \bm \Psi$ has dimension less than $t$. When for instance
$Y=\C^v$, we have $T_{\bm y} Y=\C^v$ and this condition is equivalent
to the Jacobian of $\bm \Psi$ having rank less than $t$ at $\bm y$.  A
{\em regular point} of $\bm \Psi$ is a point on $Y$ that is not
critical. {\em Critical values} are the images of critical points; the
complement of this set are the {\em regular values} (so a regular
value is not necessarily in the image of $\bm\Psi$).

Let $n,\dt,$ and $m$ be positive integers, with $m \le n$, and denote
by $\bm\Phi :\C^n \times \C^{\dt} ~ \rightarrow \C^{m}$ a mapping
defined by polynomials in $\C[\Xb,\Thetab]$, where $\Xb$,
resp.\ $\Thetab$, is a set of $n$, resp.\ $\dt$, indeterminates.  For
$\thetab$ in $\C^{\dt}$, let $\bm\Phi_{\thetab} : \C^n \rightarrow
\C^{m}$ be the induced mapping $\xb\mapsto \bm\Phi(\xb,\thetab)$.  The
transversality result we will need is the following.

\begin{proposition} [Weak transversality]\label{prop:weak_t}
  Suppose that $\bm 0$ is a regular value of $\bm\Phi$. Then there
  exists a hypersurface $\Delta \subset \C^{\dt}$ of degree at most
  \td for which, if $\thetab \in \C^{\dt}-\Delta$, then $\bm 0$ is a
  regular value of~$\bm\Phi_{\thetab}$.
\end{proposition}

Here is a simple example that shows this result at work. Consider a
squarefree $f$ in $\C[X_1,X_2]$, such that $V(f)$ is a smooth curve in
$\C^2$, and let the mapping $\bm\Phi:\C^2\times \C \to \C^2$ be defined
by $\bm\Phi(X_1,X_2,\Lambda) = (f(X_1,X_2), X_1-\Lambda)$. One checks
that the Jacobian of $\bm\Phi$ with respect to $(X_1,X_2,\Lambda)$ has
rank two at any point in $\bm\Phi^{-1}(\bm 0)$, so the assumptions of the
proposition apply. We deduce that for a generic $\lambda$ in $\C$,
that is, for all $\lambda$ in $\C$ except a finite number, the ideal
$(f(X_1,X_2), X_1-\lambda)$ is radical in $\C[X_1,X_2]$; equivalently,
$f(\lambda, X_2)$ is squarefree. We will revisit this example in
Section~\td.

The rest of the section is devoted to the proof of the proposition.
The proof of \cite[Theorem B.3]{TWT} already shows the existence of
$\Delta$; it is essentially the classical proof for smooth
mappings~\cite[Section~3.7]{demazure2000bifurcations}, written in an
algebraic context. In what follows, we revisit this proof,
establishing a bound on the degree of $\Delta$.

We will reuse the following fact proved in~\cite{TWT}.  Put $V =
\bm\Phi^{-1}(0)$ and consider the projection $\pi:(\xb, \thetab) \in
\C^n \times \C^{\dt} \mapsto \thetab \in\C^{\dt}$.  If $V$ is empty,
there is nothing to do, since all values $\thetab$ in $\C^\dt$ satisfy
the conclusion of the theorem. Thus, we assume that $V$ is not
empty. Then, the Jacobian criterion shows that $V$ is
smooth and $(n+ \dt -m)$-equidimensional.

Let $Z$ be the set of critical points of $\pi_{|V}$, and consider its
projection $\pi(Z)$ in $\C^s$. This is the set of critical values of
$\pi_{|V}$; hence, by the algebraic form of Sard's lemma (see
\cite[Theorem~3.7]{Mumford76} for irreducible $V$
and~\cite[Proposition~B.2]{TWT} for general $V$), its Zariski closure
$\overline{\pi(Z)}$ is a strict closed subset of $\C^s$. Now, the
proof of \cite[Theorem B.3]{TWT} shows that if $\thetab \in \C^{\dt}$
is not in $\overline{\pi(Z)}$, then $\bm 0$ is a critical value of
$\bm\Phi_{\thetab}$. Hence, it suffices to prove that $Z$ has degree
at most $d^{m+n}$; this will imply the same degree bound for
$\overline{\pi(Z)}$, and prove our propositon.

We first describe the set $Z$ of critical points of $\pi_{|V}$.  Let
$M$ denote the $(s+m) \times (s+n)$ matrix with entries in
$\C[\Xb,\Thetab]$ given by $M = \jac_{\Xb,\Thetab}(\pi,\bm\Phi)$, that
is,
\begin{align*}
M &= 
\bbm 
\jac_{\Xb,\Thetab}(\pi)\\
\jac_{\Xb,\Thetab}(\bm\Phi) 
\ebm 
=
\bbm 
\textbf{0}_{\dt \times n}\hspace{5mm}\textbf{I}_{\dt} \\
\jac_{(\Xb,\Thetab)}(\bm\Phi)
\ebm.
\end{align*}

\begin{lemma} 
  For $(\xb,\thetab)$ in $V$, the matrix $M$ has full rank $\dt+m$ at
  $(\xb,\thetab)$ if and only if $(\xb,\thetab)$ is a regular point of
  $\pi_{|V}.$
\end{lemma}
\begin{proof}
  Take $(\xb,\thetab)$ on $V$, and let $K(\xb,\thetab)$ be the
  Jacobian matrix $\jac_{(\Xb,\Thetab)}(\bm\Phi)$ at $(\xb,\thetab)$.
  Then, the rank of $M(\xb,\thetab)$ can be written as
  $\textup{rank}(K(\xb,\thetab)) + \textup{rank}([\textbf{0}_{\dt
      \times n}~\textbf{I}_{\dt}] \mid \ker K(\xb,\thetab))$, where
  the latter is the rank of the restriction of $[\textbf{0}_{\dt
      \times n}~\textbf{I}_{\dt}]$ to the nullspace of
  $K(\xb,\thetab)$.

  Since $V$ is smooth, $K(\xb,\thetab)$ has full rank ${\rm codim} (V)
  = m$. On the other hand, the nullspace of $K(\xb,\thetab)$ is the
  tangent space $T_{\xb,\thetab} V$, and
  $\textup{rank}([\textbf{0}_{\dt \times n}~\textbf{I}_{\dt}] \mid
  \ker K(\xb,\thetab))$ is the dimension of $\pi(T_{\xb,\thetab} V)$.
  In other words, the rank of $M(\xb,\thetab)$ is equal to
  $m+\dim(\pi(T_{\xb,\thetab} V))$; this implies the claim in the lemma.
\end{proof}

Therefore, we can characterize the set $Z$ of critical points of
$\pi_{|V}$ as those points satisfying $\bm\Phi(\xb,\thetab)=\bm 0$ and
where all minors of $M$ of order $\dt+m$ vanish. We can describe this
set using fewer equations, by discarding certain minors that are
identically zero. Let indeed $J$ denote the $m \times n$ submatrix of
the Jacobian of $\bm\Phi$ consisting of the first $n$ columns.
\begin{lemma} 
  For $(\xb,\thetab)$ in $V$, $(\xb,\thetab)$ is in $Z$ if and only if
  $J(\xb,\thetab)$ has rank less than $m$.
\end{lemma}
\begin{proof}
Notice 
%
\begin{align*}
M(\xb,\thetab)&= 
\bbm 
\textbf{0}_{\dt \times n} &\textbf{I}_{\dt} \\
J(\xb,\thetab)     &J'(\xb,\thetab)
\ebm,
\end{align*}
where $J'$ consists of the remaining columns of the Jacobian matrix of
$\bm\Phi$.  Then, the rank of the former matrix is equal to the rank
of
\begin{align*}
M(\xb,\thetab)&= 
\bbm 
\textbf{0}_{\dt \times n} &\textbf{I}_{\dt} \\
J(\xb,\thetab)     & \bm 0_{m \times s}
\ebm,
\end{align*}
and the conclusion follows.
\end{proof}
Therefore, the set $Z$ can be defined as the set of all
$(\xb,\thetab)$ in $\C^{n+\dt}$ satisfying $\bm\Phi(\xb,\thetab)=\bm
0$ and where all minors of $J$ of order $m$ vanish. Our next step is
to bound the degree of $Z$. For this, rather than considering minors
of $J$, we will rewrite the condition that $J(\xb,\thetab)$ has rank
less than $m$ as the existence of a non-trivial left kernel element.

For this, we let $\textit{\textbf{L}}=[L_1,\hdots,L_m]$ be new
variables, thought of as Lagrange multipliers, and consider the
``Lagrange polynomials'' $\mathscr{L}_1,\dots,\mathscr{L}_{n}$, with
\[
[\mathscr{L}_1 \cdots \mathscr{L}_{n}]=\textit{\textbf{L}}\cdot J(\xb,\thetab).
\] 
Denote by $\frak{Z} \subset \C^{n+\dt+m}$ the algebraic set defined by
the vanishing of $\mathscr{L}_1,\hdots,\mathscr{L}_{n},$ and
$\bm\Psi$, and by $\frak{Z}'$ the algebraic set
\[
\frak{Z}' := \overline{\frak{Z} - \{(\textit{\textbf{x}},\thetab,0,\hdots,0) \in \C^{n+\dt+m}~|~(\textit{\textbf{x}},\thetab,0\hdots,0) \in \frak{Z}\}},
\]
where the bar denotes Zariski closure (we have to remove such points,
since $L_1=\cdots=L_m=0$ is always a trivial solution to the Lagrange
equations). Finally, consider the projection
\begin{align*} 
\mu :~ \C^{n+\dt+m} &\rightarrow \C^{n+\dt}\\
(\textit{\textbf{x}},\thetab,\bm \ell)~ &\mapsto (\textit{\textbf{x}},\thetab).
\end{align*}

\begin{lemma}  
  The algebraic set $Z$ is equal to the projection $\mu(\frak{Z}')$.
\end{lemma}
\begin{proof}
  Take $(\xb,\thetab)$ in $Z$. Then, $(\xb,\thetab)$ cancels all
  polynomials $\bm \Phi$, and there exists $\bm
  \ell=(\ell_1,\dots,\ell_m)$, not identically zero, such that
  $(\xb,\thetab,\bm \ell)$ cancels the Lagrange polynomials. This
  implies that $(\xb,\thetab,\bm \ell)$ is in $\frak{Z} -
  \{(\textit{\textbf{x}}',\thetab',0,\hdots,0) \in
  \C^{n+\dt+m}~|~(\textit{\textbf{x}}',\thetab',0\hdots,0) \in
  \frak{Z}\}$, and thus in $\frak{Z}'$. This proves the inclusion
  $Z \subset \mu(\frak{Z}')$.

  Conversely, take an irreducible component $Y$ of $\frak Z'$. We prove 
  that $\mu(Y)$ is contained in $Z$.
  By construction, there exists an open dense subset $Y^o \subset Y$
  such that for any $(\xb,\thetab,\bm \ell)$ in $Y^o$, $\bm \ell$ is
  not identically zero. As a result, $(\xb,\thetab)$ is in $Z$, that
  is, $\mu(Y^o)$ is in $Z$. This implies that its Zariski closure
  $\overline{\mu(Y^o)}$ is in $Z$. Since $\mu(Y)$ is contained in
  $\overline{\mu(Y^o)}$, we have $\mu(Y) \subset Z$. Taking 
  the union over all $Y$, we get $\mu(\frak Z') \subset Z$, as claimed.
\end{proof}
%
\begin{corollary} 
  The degree of $Z$ is at most $d^{m+n}.$
\end{corollary}
\begin{proof}
  The algebraic set $\frak Z$ is defined by $m+n$ equations, all of
  them having degree at most $d$. It follows from B\'ezout's
  Theorem~\cite{H} that $\deg(\frak{Z}) \leq d^{m+n}$, and the same
  upper bound holds for $\deg(\frak{Z}')$. Applying the projection
  $\mu$ yields the result.
\end{proof}

@@

It will sometimes be useful for us to let the matrix $A$ represent new
indeterminates: let $\A$ be an $n \times n$ matrix of new
indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in
\mathbb{Q}(\A_{i,j})[\textbf{X}]$ as $f(\A\textbf{X})$ and
$V^\A=V(f^\A).$ It follows from Proposition 4.2 that for a generic $A
\in \C^{n^2},$ $0$ is a regular value of $\Phi_{A}$.  The proof of
\cite[Theorem B.3]{NO}, which shows the existence of $\Delta \subset
\C^{n^2},$ applies \textit{Sard's lemma}, which shows that the
critical values of a polynomial mapping are contained in a
hypersurface. It follows from Sard's lemma that for the matrix of
indeterminants $\A,$ $0$ is a regular value of $\Phi_{\A}.$ Otherwise
it could not be that $0$ is a regular value of $\Phi_A,$ for a generic
$A\in \C^{n^2}.$ Therefore,

@@

has full rank for all $\xb \in V(f^{\A})$, and thus $f^{\A}$satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$. We will rely on this fact in Section 5. 

@@

\section{Proof of $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$}

@@

If the Jacobian of $\fb$ has full rank at all $x \in V(f) \cap \mathbb{R}^n,$ then it follows by the Jacobian criterion [4, Theorem 16.19] that $f$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$. 

@@

There exists a Zariski closed set $\Delta \subset \C^{n^2}$ of degree at most $D^n$ with the property that, if $A \in \C^{n^2} - \Delta$ then for each $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$.

@@

has full rank for all $\xb \in V(f^A)$. Then, by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$.

@@

has full rank for all $\xb \in V(f^A)$ ($\cap \R^n$?). Therefore, by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$, for each $i \in \{1,\hdots,n\}.$

@@

\section{Proof of $\textbf{H}_i'(4),\textbf{H}_i'(5)$ and $\textbf{H}_i'(6)$}

@@

\section{Proof of $\textbf{H}_i'(7)$}
